---
title: "Homework2"
author: "Miles Tweed"
date: "1/24/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# 3. Practicing with the Titanic data

## 3.1 Find the overall proportion of males and females who survived the sinking of the Titanic.
(Ignore whether child or adult.)
```{r}
mydata <- as.data.frame(Titanic)

mydata %>% 
  count(Sex, Survived, wt = Freq) %>% 
  group_by(Sex) %>%  
  mutate(Total=sum(n), Proportion = n/Total) %>% 
  filter(Survived == 'Yes') %>% 
  rename(Count = 'n')
```

## 3.2 Plot the overall proportion of survival in a bar graph, by gender
```{r}

mydata %>% 
  count(Sex, Survived, wt = Freq) %>% 
  group_by(Sex) %>%  
  mutate(tot=sum(n), prop = n/tot) %>%
  filter(Survived == 'Yes') %>% 
  ggplot(aes(x = Sex, y = prop, fill = Sex)) + 
  geom_bar(stat = 'identity', width = 0.7) +
  labs(title = 'Titanic Survival By Gender') + 
  ylab('Proportion Survived') +
  theme(legend.position = 'none')
```

## 3.3 Find the overall proportion of males and females who survived the sinking of the Titanic, **by class booked**.
(Ignore whether child or adult.)
```{r}
# because we want props of survival by sex and class, we count by sex, class and survival

mydata %>% 
  count(Sex, Class, Survived, wt=Freq) %>% 
  group_by(Sex, Class) %>% 
  mutate(Total = sum(n), Proportion = n/Total) %>% 
  filter(Survived=='Yes') %>% 
  rename(Count = 'n')


```


## 3.4 Plot the survival proportions by class
There are a couple of options: For each class, we could compare males vs. females:
```{r}
mydata %>% 
  count(Sex, Class, Survived, wt=Freq) %>% 
  group_by(Sex, Class) %>% 
  mutate(Total = sum(n), Proportion = n/Total) %>% 
  filter(Survived=='Yes') %>% 
  rename(Count = 'n') %>% 
  ggplot(aes(x = Class, y = Proportion, fill = Sex)) + 
  geom_bar(stat = 'identity', position = 'dodge') +
  labs(title = 'Proportion of Survival by class')

```

Or, we could compare the proportions surviving in each class, separately for males and females:
```{r}
mydata %>% 
  count(Sex, Class, Survived, wt=Freq) %>% 
  group_by(Sex, Class) %>% 
  mutate(Total = sum(n), Proportion = n/Total) %>% 
  filter(Survived=='Yes') %>% 
  rename(Count = 'n') %>% 
  ggplot(aes(x = Sex, y = Proportion, fill = Class)) + 
  geom_bar(stat = 'identity', position = 'dodge') +
  labs(title = 'Proportion of Survival by class')
```

## 3.5 Same as 3.3, but now also adjust for whether child or adult (the `age` variable).
So, find the proportion of passengers who survived, by sex, class and whether child or adult.
```{r}
mydata %>% 
  count(Sex, Class, Survived, Age, wt=Freq) %>% 
  group_by(Sex, Class, Age) %>% 
  mutate(Total = sum(n), Proportion = n/Total) %>% 
  filter(Survived == 'Yes')

```

Or, more succinctly, using `add_count`:
```{r}
mydata %>% 
  add_count(Class,Sex,Age, wt = Freq) %>% 
  group_by(Class, Sex, Age) %>% 
  mutate(Proportion = Freq/n) %>% 
  filter(Survived == 'Yes')
```


## 3.6 Plot the proportions when adjusting for all three variables (Sex, Class, Age): `facet_grid`
When comparing the proportions, we are now adusting for three variables: Sex, Class and Age. To get all these in one plot, we will assign one variable to be a row variable, one variable to be a column variable, and a third to be the variable that is displayed on the x-axis. Naturally, there are many possibilities. Let's first construct a graph that, similar to the second to last above, has sex on the x-axis, class as the column variable and age as the row variable. We assign the row and column variable in `facet_grid`, which we now use instead of `facet_wrap`:
```{r}
mydata %>% 
  add_count(Class,Sex,Age, wt = Freq) %>% 
  group_by(Class, Sex, Age) %>% 
  mutate(Proportion = Freq/n) %>% 
  filter(Survived == 'Yes', !is.na(Proportion)) %>% 
  ggplot(aes(x = Sex, y = Proportion, fill = Sex)) + 
  geom_bar(stat = 'identity') + 
  facet_grid(rows = vars(Age), cols = vars(Class)) + 
  theme(legend.position = 'none')
```

If you want the label of the column variable appear on the left side instead of on the right-side, use the `switch='y'` options:
```{r}
mydata %>% 
  add_count(Class,Sex,Age, wt = Freq) %>% 
  group_by(Class, Sex, Age) %>% 
  mutate(Proportion = Freq/n) %>% 
  filter(Survived == 'Yes', !is.na(Proportion)) %>% 
  ggplot(aes(x = Sex, y = Proportion, fill = Sex)) + 
  geom_bar(stat = 'identity') + 
  facet_grid(rows = vars(Age), cols = vars(Class), switch = 'y') + 
  theme(legend.position = 'none')
```

Try other versions of this plot, by changing what you assign for row, column and x-axis variable!

```{r}
mydata %>% 
  add_count(Class,Sex,Age, wt = Freq) %>% 
  group_by(Class, Sex, Age) %>% 
  mutate(Proportion = Freq/n) %>% 
  filter(Survived == 'Yes', !is.na(Proportion)) %>% 
  ggplot(aes(x = Class, y = Proportion, fill = Class)) + 
  geom_bar(stat = 'identity') + 
  facet_grid(rows = vars(Age), cols = vars(Sex)) + 
  theme(legend.position = 'none')
```

