---
title: "Homework # ..., My Name."
output:
  pdf_document: default
  html_document: default
---

# This one contains solutions to some questions from both HW 1 \& 2, simply for illustration of expectations for both (even though you haven't gotten the HW2 yet).

## Problem 1 (make sure to include code with outputs)

1. I will study the political affiliations (**keep in mind - you can't pick that variable**), which takes on three categories: democrat ("d"), independent ("i") and republican ("r"):

```{r, fig.height=3.5}
fl_data <- read.csv("http://sites.williams.edu/bklingen/files/2015/07/fl_student_survey.csv") # You file path will probably be different
table(fl_data$political_affiliation)
barplot(table(fl_data$political_affiliation))
```

Most students had independent political affiliation (24), second most - democratic (21), and the least - republican (15).

2. ...

## Problem 2

**1.7** 

a) 3077 people who responded;

b) All adults in the United States;

c) 23%.

**1.8** ...

**1.9** 

a) A graduate student;

b) A few students from a particular graduate program;

c) All students from that graduate program.

**1.10** ...

**2.3**

**2.6**

**2.10** 

**2.21**  

a) Skewed to the right, because of a few expensive upscale houses.

b) Skewed to the right, because of a few faculty tending to overdraw frequently.

c) Symmetric, due to most being around average, with several over- and underachievers.

d) Symmetric, just like with most physical characteristics: most are around average, with several really short and really tall girls.

**2.34** 

**2.37**

```{r}
miles <- c(0,0,4,0,0,0,10,0,6,0)
mean(miles)
sort(miles)

```
$$Mean = (0+0+4+0+0+0+10+0+6+0)/10 = 2$$
$$ Median = \{5th + 6th \ element\}/2 = (0+0)/2=0$$
$$ Mode = \{Most \ frequent \ value\} = 0$$
Interpretations:

  - *Mean*: On a typical day, company employees average 2 miles of travel using public transportation.
  - *Median*: On a typical day, median # of miles traveled by employees via public transportation is $(0+0)/2 = 0$ (average of 5th and 6th employees).
  - *Mode*: Most of the employees travel 0 miles using public transportation on a typical day.

...

**2.46**

**2.54**

**2.62**

**2.66**

**2.67**

**2.76**
