---
title: "Homework7"
author: "Miles Tweed"
date: "1/29/2021"
output: pdf_document
---

Please complete points 4 and 5

Fit a simple linear regression model for the Ames house price data set (on Canvas), using the price of the house (in $1000) as the response variable and the size as the explanatory variable.
  
Compute the standardized (or better yet: studentized) residuals. (rstudent(fit) returns studentized residuals, where fit = lm(y~x).)
    
Plot a (faint) histogram of the residuals, with a smooth density estimate and a standard normal distribution overlaid.
    
Obtain a P-P plot of the studentized residuals.

Obtain a Q-Q plot of the studentized residuals and briefly explain.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
house <- read_csv('/home/mtweed/Documents/New_College/year_2/Data_viz/Data/traindata_housepricesAmes.csv')

house <-  house %>% mutate(SF = `1stFlrSF` + `2ndFlrSF`)

model <- lm(SalePrice/1000 ~ SF, data = house)

house <- house %>% mutate(stresid = rstudent(model))

```


```{r}
house %>% ggplot() + 
  geom_histogram(aes(x = stresid, y = ..density..), 
                 binwidth = 0.66, 
                 alpha = 0.25,
                 color = 'gray91',
                 fill = 'seashell4'
                 ) + 
  geom_density(aes(x = stresid, color = 'Residuals')) + 
  stat_function(aes(color = "Normal"),
                fun = dnorm, 
                args = list(mean = mean(house$stresid), sd = sd(house$stresid)),
                size=0.9) + 
  labs(title="Residual Density", x="", y="", color = "Distribution") +
  theme(axis.line.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.text.y=element_blank(),
        axis.title.y = element_blank()) +
  theme_minimal() + 
  ggsave('residualDensity.png', width = 150, height = 75, units = 'mm')

```

```{r fig.height=2, fig.width=3}
mysample <- house$stresid
F.x <- ecdf(mysample)
yvar <- F.x(mysample)
xvar <- pnorm(mysample, mean=mean(mysample), sd=sd(mysample))
df <- data.frame(x=xvar, y=yvar)

ggplot(data=df, aes(x=x, y=y)) + 
  geom_point(col="blue") + 
  geom_abline(intercept=0, slope=1, color="red") +
  coord_cartesian(xlim=c(0,1), ylim=c(0,1)) +
  theme_minimal() +
  labs(y="Residual Cum. Prob.", x="Normal Cum. Prob.", title="P-P Plot: Normal") + 
  ggsave('PPnorm.png', width = 150, height = 75, units = 'mm')
```

```{r fig.height=2, fig.width=3}
mysample <- house$stresid
F.x <- ecdf(mysample)
yvar <- F.x(mysample)
xvar <- pexp(mysample)#, mean=mean(mysample), sd=sd(mysample))
df <- data.frame(x=xvar, y=yvar)

ggplot(data=df, aes(x=x, y=y)) + 
  geom_point(col="blue") + 
  geom_abline(intercept=0, slope=1, color="red") +
  coord_cartesian(xlim=c(0,1), ylim=c(0,1)) +
  theme_minimal() +
  labs(y="Residual Cum. Prob.", x="Exponential Cum. Prob.", title="P-P Plot: Exponential")+ 
  ggsave('PPexp.png', width = 150, height = 75, units = 'mm')
```



```{r fig.height=2, fig.width=3}
quantile(mysample)
ggplot(data=data.frame(x=mysample), aes(sample=mysample)) + 
  geom_qq(dparams=list(mean=mean(mysample), sd=sd(mysample)), color="blue") +
  geom_qq_line(dparams=list(mean=mean(mysample), sd=sd(mysample)), color="red") +
  theme_minimal() +
  labs(y="Residual Quantile", x="Theoretical Quantile", title="Q-Q Plot") 
  ggsave('QQnorm.png', width = 150, height = 75, units = 'mm')
```


